### 2018-06-10 MSc_Marko_Docevski_tekst.tex
Updates from ShareLaTeX

\cite{Liang2017} While the method we develop is capable of modeling any multi-part music, we limit the scope of this work to Bach’s chorales because: they provide a relatively large corpus, by a single composer, are well understood by music theorists, and are routinely used in the teaching of music theory. They intro- duce the JSB Chorales dataset which has since become a standard benchmark routinely used to evaluate the per- formance of generative models on polyphonic music mod- elling 
While BachBot also utilizes a LSTM for capturing long range dependencies, BachBot uses a softmax distri- bution rather than a DBN to parameterize the probability distribution and hence does not require Monte Carlo sam- pling at each time step of training and inference. 
Model Architecture, Training, and Sampling
We use a RNN with LSTM memory cells and the following hyperparameters: 1. num layers – the number of memory cell layers
2. rnn size – the number of hidden units per mem- ory cell (i.e. hidden state dimension)
3. wordvec – dimension of vector embeddings
4. seq length – number of frames before truncating back-propagation through time (BPTT) gradient
5. dropout – the dropout probability
Our model first embeds the inputs xt into a wordvec- dimensional vector-space, compressing the dimensionality
down from |V | ≈ 140 to wordvec dimensions. Next, num layers layers of memory cells followed by batch normalization [28] and dropout [26] with dropout probability dropout are stacked. The outputs y(num layers) are t followed by a fully-connected layer mapping to |V | = 108 units, which are passed through a softmax to yield a predictive distribution P(xt+1|ht−1,xt): the probability distribution over the next token xt+1 given the current token xt and the previous RNN memory cell state ht−1. Models were trained using the Adam optimizer [29] with a minibatch size of 50 and an initial learning rate
of 2 × 10−3 decayed by 0.5 every 5 epochs. The back- propagation through time gradients were clipped at ±5.0 [32] and truncated after seq length frames
We minimize cross-entropy loss between the predicted distributions P(xt+1|xt,ht−1) and the actual target dis- tribution δxt+1. During training, the correct token xt+1 is treated as the model output even if the most likely predic- tion argmaxP(xt+1|ht,xt) differs. 
During inference, we per- form ancestral sampling and reuse the actual token ˆxt sampled from P(xt|ht−1,xt−1) to compute P(xt+1|ht,xt) for sampling ˆ
xt+1. Unlike MCMC, which requires running multiple iterations to obtain a single sample, ancestral sampling requires only a single forward pass.
We encode the scores into sequences of tokens amenable for sequential processing by recurrent neural networks (RNNs). We limit the symbolic representation to pitch and rhythm. we avoid explicitly encoding music- theoretic concepts such as motifs, phrases, and chords / inversions, instead tasking the model to learn musically meaningful features with minimal prior knowledge
Our encoding represents polyphonic scores with sixteenth-note frames, encoding duration implicitly by the number of frames processed. Such an encoding requires the network to leverage memory to account for longer du- rations notes, a counting and timing task which LSTM is known to be capable of [19]. Consecutive frames are sep- arated by a unique delimiter
Within each frame, we represent individual notes rather than entire chords, reducing the vocabulary size from O(1284) down to O(128). Prior work modeling charac- ters versus words in language modeling tasks suggests that this has negligible impact [22]. Each frame consists of
four (Soprano, Alto, Tenor, and Bass) ?Pitch,Tie? tu- ples where Pitch ∈ {0, 1, · · · , 127} represents the MIDI pitch of a note and Tie ∈ {True,False} distinguishes whether a note is tied with a note at the same pitch from the previous frame or is articulated at the current timestep. We order notes within a frame in descending MIDI pitch and neglects crossing voices;
Within each frame, we represent individual notes rather
than entire chords, reducing the vocabulary size from O(1284) down to O(128). Prior work modeling charac- ters versus words in language modeling tasks suggests that this has negligible impact [22]. Each frame consists of
four (Soprano, Alto, Tenor, and Bass) ?Pitch,Tie? tu- ples where Pitch ∈ {0, 1, · · · , 127} represents the MIDI pitch of a note and Tie ∈ {True,False} distinguishes whether a note is tied with a note at the same pitch from the
previous frame or is articulated at the current timestep. We order notes within a frame in descending MIDI pitch and neglects crossing voices;
EXPERIMENTS 3.1 Sequence Modelling
With the BachBot model, we performed a grid search through the parameter grid in table 1 and found num layers = 3, rnn size = 256, wordvec = 32, seq length = 128 dropout = 0.3 achieves the low- est cross-entropy loss of 0.477 bits on a 10percent held-out val- idation corpus.
We extend this harmonization task to the completion of chorales for a wider number and type of given parts. Let
x(1:T) be a sequence of tokens representing an encoded musical score, {1, 2, . . . ,T} a multi-index, and sup- pose ? correspond to some fixed token values to be harmonized (e.g. a provided Soprano line). 
Musical Discrimination Test
To measure BachBot’s success in this task, we devel- oped a publicly accessible musical discrimination test at bachbot.com. Unlike prior studies which leverage paid services like Amazon MTurk for human feedback [35], we offered no such incentive and promoted the study only through social media. Participants were first surveyed for their age group and prior music experience (fig. 3a). Next, they are presented five discrimination tasks which presented two audio tracks (an original Bach composition and a synthetic composition by BachBot) and ask them to identify the Bach original. Each audio track contains an entire composition from start to end. The music score for the audio was not provided. Participants were granted an unlimited amount of time and allowed to replay each track an arbitrary number of times. Participants could only see the next question after submit- ting the current one and were not allowed to modify their responses after submitting. The five questions comprised of three harmonizations (S/A/T/B, one AT, one ATB), and two original compositions. To construct the questions, harmonizations were paired along with the original Bach chorales the fixed parts were taken from. No such direct comparison is possible for the SATB case, so these synthetic compositions were paired with a randomly selected Bach chorale in a some- what different comparative listening task. Harmonizations were synthesized by extracting part(s) from a randomly se- lected Bach chorale and filling in the remaining parts of the composition using the method previously described in sec- tion 2.4. Original compositions (questions labelled SATB) were generated by providing a START symbol followed by ancestral sampling as previously described in section 2.3 until an END symbol is reached. The final audio provided in the questions were obtained by rendering the composi- tions using the Piano instrument from the Fluid R3 GM SoundFont.
We only considered the first response per IP address of participants who had played both choices in every question at least once and completed all five questions. This totaled 2336 participants at the time of writing, making our study one of the largest subjective listening evaluation of an au- tomatic composition system to date.
An informal analysis sug- gests that while some neurons are ambiguous to interpreta- tion, other neurons correlate significantly with recognized music-theoretic objects, particularly chords (see fig. 4). To our knowledge, this is the first reported evidence for an LSTM optimized for automatic composition learning music-theoretic concepts without explicit prior informa- tion. This invites a follow-up study testing the statistical significance of these observations.

@@

Користењето на јазични модели на ниво на бука овозможува едноставно моделирање на монофона музика, или музика која во секој момент има највеќе една активна нова. При моделирање на акорди со пиано лента се појавува проблем на енумерирање сите можни конфигурации на ноти што може да се појават, пр. за гитара има грубо $2*24^6$ можни конфигурации на ноти. Еден пристап за намалување на проблемите што доаѓаат од зголемена димензионалност може да се види во \cite{Boulanger-Lewandowski2012, Boulanger-Lewandowski2014, Goel2014}, каде што авторите користат хибридна архитектура, изграден од рекурентни невронски слоеви во различни комбинации со енергетски модели. Идејата е да се искористат особеностите на енергетските модели, како што се RBM (Restricted Boltzmann Machines - Ограничени Болзманови Машини) и NADE (Neural Autoregressive Distribution Estimator), да моделираат веројатностни дистрибуции, кон моделирање на истовремено појавување на нотите во акорди. Од \cite{Boulanger-Lewandowski2012, Boulanger-Lewandowski2014} произлегуваат моделите: RTRBM (Recurrent Temporal RBM - Рекурентна Темпорална Ограничена Болзманова Машина), RNN-RBM, RNN-NADE, од кои RNN-NADE се покажал како наједноставен за тренирање и со најдобри генеративни резултати и најекспресивен, иако од енергетските модели NADE има релативно помала експресивна моќ. Моделот може да се разбере како низа од условени енергетски слоеви, по еден за секој временски чекор што го обработува мрежата, со излез во детерминистички рекурентен невронски слој. Наголем проблем на овој тип на модели се покажала осетливоста на иницијална состојба, што многу ја потенцира важноста на пред-тренирање при нивна употреба. Моделот се покажал успешен во моделирање на локални зависности, хармониски правила и кратки мелодиски линии, меѓутоа не може да моделира долговременски зависности.

\cite{Goel2014} Elman networs were used for chaotic inspiration generated music, however they used chaotic inspiration as an exernal input instead of real stochasticity. Deep belief networks DBNs are stacked RBMs that have been trained gredily. They are graphical models which learn to extract a deep hierarchical representation of the training data. They model the joint distribution between observed vector x and the ℓ hidden layers hk. DBNs can be constructed out of RBM layers, with greedy layer-wise training. It starts from a input layer and an RBM layer on top which learns to model the input layer. When it's trained another layer is added on top, that learns to model the transformed data (samples or mean activations). As many layers as needed can be added hierachicaly. The Recurrent Temporal Restricted Boltzmann Machine (RTRBM) is a sequence of conditional RBMs (one at each time instant) whose parameters are time-dependent and depend on the sequence history at time t, denoted by , where u(t) is the mean-field value of h(t), The RTRBM is formally defined by its joint probability distribution. The RTRBM can be thought of as a sequence of conditional RBMs whose parameters are the output of a deterministic RNN , with the constraint that the hidden units must describe the conditional distributions and convey temporal information for sequence generation. The use of a single RBM layer greatly constricts the expressive power of the model as a whole. To increase the expressive power of and RTRBM, the RBM structure can be replaced with a more expressive DBN. This comination they call the RNN-DBN. RNN-DBN training algorithm. Their RNN-DN architecture consists of 2 DBN layers, each having 150 binary units and 150 binary units in the RNN layer. The visible layer has 88 binary units, corresponding to the full range of piano A0 to C8. They didn't use any preprocessing techniques, only raw data is fed to the net. Model is evaluated qualitatatively with log-likelihood as a perf measure. No preprocessing was applied. Using any of the techniquest with transposition to common tonality, transposition to different keys, normalizating the tempo in beats per minute could improve the results.Pre-training could improve the network learning process. Optimization techniques like gradient clipping, Nesterov momentum, NADE for conditional density estimations could also improve perf.

\cite{Hadjeres2016} In this paper we introduce DeepBach, a dependency net- work (Heckerman et al., 2000) capable of producing musi- cally convincing four-part chorales in the style of Bach by using a Gibbs-like sampling procedure. NOTES AND VOICES We use MIDI pitches to encode notes and choose to model voices separately. We consider that only one note can be sung at a given time and discard chorales with voice divi- sions. Since Bach chorales only contain simple time signatures, we discretize time with sixteenth notes, which means that each beat is subdivided into four equal parts. Since there is no smaller subdivision in Bach chorales, there is no loss of information in this process. RHYTHM We choose to model rhythm by simply adding a hold sym- bol “ ” coding whether or not the preceding note is held to the list of existing notes. This representation is thus unam- biguous, compact and well-suited to our sampling method (see
CHORALE We represent a chorale as a couple (V,M) (1) composed of voices and metadata. For Bach chorales, V is a list of 4 voices Vi for i ∈ [4] (soprano, alto, tenor and bass) andMa collection of metadata lists (F and S). Our choices are very general and do not involve expert knowledge about harmony or scales but are only mere ob-
servations of the corpus. The list S acts as a metronome. The list F is added since fermatas in Bach chorales indi- cate the end of each musical phrase. The use of fermata to this end is a specificity of Bach chorales that we want to take advantage of
For accurate predictions and in order to take into account the sequential aspect of the data, each classifier is mod- eled using four neural networks: two Deep Recurrent Neu- ral Networks (Pascanu et al., 2013), one summing up past information and another summing up information coming from the future together with a non-recurrent neural net- work for notes occurring at the same time. Only the last output from the uppermost RNN layer is kept. These three outputs are then merged and passed as the input of a fourth neural network whose output is pi(Vt
i |V\i,t,M, θ). Figure 4 shows a graphical representation for one of these models. Details are provided in Sect. 2.4. These choices of architecture somehow match real compositional practice on Bach chorales. Indeed, when reharmonizing a given melody, it is often simpler to start from the cadence and write music “backwards.” Generation 2.3.1. ALGORITHM Generation in dependency networks is performed using the pseudo-Gibbs sampling procedure. This Markov Chain Monte Carlo (MCMC) algorithm is described in Alg.1. It is similar to the classical Gibbs sampling procedure (Geman & Geman, 1984) on the difference that the conditional dis- tributions are potentially incompatible (Chen & Ip, 2015). This means that the conditional distributions of Eq. (2) do not necessarily comes from a joint distribution p(V) and that the theoretical guarantees that the MCMC converges to this stationary joint distribution vanish. We experimen- tally verified that it was indeed the case by checking that the Markov Chain of Alg.1 violatesKolmogorov’s criterion (Kelly, 2011): it is thus not reversible and cannot converge to a joint distribution whose conditional distributions match the ones used for sampling. However, this Markov chain converges to another station- ary distribution and applications on real data demonstrated that this method yielded accurate joint probabilities, espe- cially when the inconsistent probability distributions are learned from data (Heckerman et al., 2000). Furthermore, nonreversible MCMC algorithms can in particular cases be better at sampling that reversible Markov Chains (Vucelja, 2014). 2.3.2. We emphasize on this section the importance of our partic- ular choice of data representation with respect to our sam- pling procedure. The fact that we obtain great results using pseudo-Gibbs sampling relies exclusively on our choice to integrate the hold symbol into the list of notes. Indeed, Gibbs sampling fails to sample the true joint dis-
tribution p(V|M, θ) when variables are highly correlated, creating isolated regions of high probability states in which theMCMCchain can be trapped. However, many data rep- resentations used in music modeling such as
• the piano-roll representation,
• the couple (pitch, articulation) representation where articulation is a Boolean value indicating whether or not the note is played or held,
tend to make the musical data suffer from this drawback.
As an example, in the piano-roll representation, a long note is represented as the repetition of the same value over many variables. In order to only change its pitch, one needs to change simultaneously a large number of variables (which is exponentially rare) while this is achievable with only one variable change with our representation. After removing chorales with instrumental parts and chorales containing parts with two simultaneous notes (bass parts sometimes divide for the last chord), we ended up with 352 pieces. Contrary to other approaches which transpose all chorales to the same key (usually in C major or A minor), we choose to augment our dataset by adding all chorale transpositions which fit within the vocal ranges defined by the initial corpus.
Discrimination Test: “Bach or Computer” experiment
Subjects were presented series of only one musical extract together with the binary choice “Bach” or “Computer”. Fig. 5 shows howthe votes are distributed depending on the level of musical expertise of the subjects for each model. For this experiment, 1272 people took this test, 261 with musical expertise 1, 646 with musical expertise 2 and 365 with musical expertise 3.
Subjects were asked to give information about their musical expertise. They could choose what category fits them best between:
1. I seldom listen to classical music
2. Music lover or musician
3. Student in music composition or professional musi- cian.
Interactive composition 4.1. Description
We developed a plugin on top of the MuseScore music editor allowing a user to call DeepBach on any rectangu- lar region. Even if the interface is minimal (see Fig.7), the possibilities are numerous: we can generate a chorale from scratch, reharmonize a melody and regenerate a given chord, bar or part. We believe that this interplay between a user and the system can boost creativity and can interest a wide range of audience.

@@

@@

>>>>

Коралите на Јохан Себастијан Бах претставуваат интересен проблем за моделирање. Се стостојат од 4 гласна полифонија, алто, тенор, сопрано и бас монофонични мелодиски линии. Може да се моделираат на сличен начин како акордите, со тоа што зависностите помеѓу гласовите не е иста како меѓу нотите во акордите. Исто така уникатните комбинации од ноти по гласови поретко се јавуваат одошто кај акордите, бидејќи акордите се стандардни конструктивни елементи во музиката. Ова податочно множество е главен фокус на трудовите \cite{Liang2017, Hadjeres2016}. Во \cite{Liang2017} е искористен релативно еднсотавна архитектура, составена од повеќе рекурентни LSTM слоеви, со embedding слој после влезниот инспириран од word2vec\cite{Herremans2017}, со едноставен софтмакс активациски слој. Бидејќи ритамот во целото податочно множество е 4/4, музиката е поделена на еднакво долги рамки со времетраење 1/16-ка, во која 4те гласа се претставени со субсеквенца од поединечни ноти, а рамките меѓусебно се поделени со посебен симбол за разграничување. Со тоа се продолжува ја намалуваат комплексноста на проблемот за предвидување од $O(128^4)$ на $O(128)$, но ја издолжуваат секвенцата со фактор 5. Со алгоритам оптимизација на хиперпараметри GridSearch ги оптимизирале следните парамерти: број на рекурентни слоеви, големина на рекурентните слоеви, големина на embedding слојот, должина на секвенца што се користи за TBPT (анг. Truncated Back Propagation Through Time, Скратена повратна инфомација назад низ времето), процент за dropout (техника за нормализација, толкувано губиток на излезна информација). Во трудот \cite{Hadjeres2016} авторите користат посебни модели за секој од гласовите, и на крај ги сумираат нивните резултати. Ова влече одредена независност помеѓу гласовите, што во реалност секако не е точна, гласовите се функционално зависни помеѓусебе. За секој глас имаа модел составен од две рекурентни длабоки невронски мрежи, од кои едната го обработува времето наназад, а другата нанапред, и една целосно поврзана невронска мрежа која ги учи истровремено појавените ноти. Овие три подмрежи работат во паралела за секој временски чекор, и излезите од сите три се користат како влез во финална општа невронска мрежа која ги влече заклучоците за секој временски чекор. За генерирање користат псевдо-Гибсово семплирање (варијанта на Монте карло Маркови синџири), со кое влечат примероци од добиените излези од 4те подмодели за секој глас. Податочното множество го имаат транспонирано во сите можни клучеви наместо стандардизирање кон еден клуч. И во двата труда се прикажани процеси за субјективна оцена на резултатите со онлајн тестови каде што на корисниците им се пушта музички исечок и треба да одлучат дали е генерирана музика или оригинална композиција на Бах. Во двата случаја најпозитивни резултати покажале кога моделите ги користеле за рехармонизација на мелодија врз основа на 2 или 3 постоечки гласа.

@@

Користењето на јазични модели на ниво на бука овозможува едноставно моделирање на монофона музика, или музика која во секој момент има највеќе една активна нова. При моделирање на акорди со пиано лента се појавува проблем на енумерирање сите можни конфигурации на ноти што може да се појават, пр. за гитара има грубо $2*24^6$ можни конфигурации на ноти. Еден пристап за намалување на проблемите што доаѓаат од зголемена димензионалност може да се види во \cite{Boulanger-Lewandowski2012, Boulanger-Lewandowski2014, Goel2014}, каде што авторите користат хибридна архитектура, изграден од рекурентни невронски слоеви во различни комбинации со енергетски модели. Идејата е да се искористат особеностите на енергетските модели, како што се RBM (Restricted Boltzmann Machines - Ограничени Болзманови Машини) и NADE (Neural Autoregressive Distribution Estimator), да моделираат веројатностни дистрибуции, кон моделирање на истовремено појавување на нотите во акорди. Од \cite{Boulanger-Lewandowski2012, Boulanger-Lewandowski2014} произлегуваат моделите: RTRBM (Recurrent Temporal RBM - Рекурентна Темпорална Ограничена Болзманова Машина), RNN-RBM, RNN-NADE, од кои RNN-NADE се покажал како наједноставен за тренирање и со најдобри генеративни резултати и најекспресивен, иако од енергетските модели NADE има релативно помала експресивна моќ. Моделот може да се разбере како низа од условени енергетски слоеви, по еден за секој временски чекор што го обработува мрежата, со излез во детерминистички рекурентен невронски слој. Наголем проблем на овој тип на модели се покажала осетливоста на иницијална состојба, што многу ја потенцира важноста на пред-тренирање при нивна употреба. Моделот се покажал успешен во моделирање на локални зависности, хармониски правила и кратки мелодиски линии, меѓутоа не може да моделира долговременски зависности. Во \cite{Goel2014} авторите користат пософистицирана варијанта на RNN-RBM наречена RNN-DBN, во која секвенците од RBM слоеви се заменети со секвенци од длабоки подмрежи изградени од повеќе хиерахиски RBM слоеви. Не користат никакви техники за иницијализација на енергетските слоеви како во \cite{Boulanger-Lewandowski2012, Boulanger-Lewandowski2014, Goel2014} ниту за оптимизирање на работата на моделот.

@@

\section{Gibbs sampling / Гибсово семплирање}

\cite{Hadjeres2016} ALGORITHM Generation in dependency networks is performed using the pseudo-Gibbs sampling procedure. This Markov Chain Monte Carlo (MCMC) algorithm is described in Alg.1. It is similar to the classical Gibbs sampling procedure (Geman & Geman, 1984) on the difference that the conditional dis- tributions are potentially incompatible (Chen & Ip, 2015). This means that the conditional distributions of Eq. (2) do not necessarily comes from a joint distribution p(V) and that the theoretical guarantees that the MCMC converges to this stationary joint distribution vanish. We experimen- tally verified that it was indeed the case by checking that the Markov Chain of Alg.1 violatesKolmogorov’s criterion (Kelly, 2011): it is thus not reversible and cannot converge to a joint distribution whose conditional distributions match the ones used for sampling. However, this Markov chain converges to another station- ary distribution and applications on real data demonstrated that this method yielded accurate joint probabilities, espe- cially when the inconsistent probability distributions are learned from data (Heckerman et al., 2000). Furthermore, nonreversible MCMC algorithms can in particular cases be better at sampling that reversible Markov Chains (Vucelja, 2014). 2.3.2. We emphasize on this section the importance of our partic- ular choice of data representation with respect to our sam- pling procedure. The fact that we obtain great results using pseudo-Gibbs sampling relies exclusively on our choice to integrate the hold symbol into the list of notes. Indeed, Gibbs sampling fails to sample the true joint dis-
tribution p(V|M, θ) when variables are highly correlated, creating isolated regions of high probability states in which theMCMCchain can be trapped. However, many data rep- resentations used in music modeling such as
• the piano-roll representation,
• the couple (pitch, articulation) representation where articulation is a Boolean value indicating whether or not the note is played or held,
tend to make the musical data suffer from this drawback.
As an example, in the piano-roll representation, a long note is represented as the repetition of the same value over many variables. In order to only change its pitch, one needs to change simultaneously a large number of variables (which is exponentially rare) while this is achievable with only one variable change with our representation.

@@

\cite{Hadjeres2016} Discrimination Test: “Bach or Computer” experiment
Subjects were presented series of only one musical extract together with the binary choice “Bach” or “Computer”. Fig. 5 shows howthe votes are distributed depending on the level of musical expertise of the subjects for each model. For this experiment, 1272 people took this test, 261 with musical expertise 1, 646 with musical expertise 2 and 365 with musical expertise 3.
Subjects were asked to give information about their musical expertise. They could choose what category fits them best between:
1. I seldom listen to classical music
2. Music lover or musician
3. Student in music composition or professional musi- cian.
Interactive composition 4.1. Description
We developed a plugin on top of the MuseScore music editor allowing a user to call DeepBach on any rectangu- lar region. Even if the interface is minimal (see Fig.7), the possibilities are numerous: we can generate a chorale from scratch, reharmonize a melody and regenerate a given chord, bar or part. We believe that this interplay between a user and the system can boost creativity and can interest a wide range of audience.

\cite{Liang2017} 
To measure BachBot’s success in this task, we devel- oped a publicly accessible musical discrimination test at bachbot.com. Unlike prior studies which leverage paid services like Amazon MTurk for human feedback [35], we offered no such incentive and promoted the study only through social media. Participants were first surveyed for their age group and prior music experience (fig. 3a). Next, they are presented five discrimination tasks which presented two audio tracks (an original Bach composition and a synthetic composition by BachBot) and ask them to identify the Bach original. Each audio track contains an entire composition from start to end. The music score for the audio was not provided. Participants were granted an unlimited amount of time and allowed to replay each track an arbitrary number of times. Participants could only see the next question after submit- ting the current one and were not allowed to modify their responses after submitting. The five questions comprised of three harmonizations (S/A/T/B, one AT, one ATB), and two original compositions. To construct the questions, harmonizations were paired along with the original Bach chorales the fixed parts were taken from. No such direct comparison is possible for the SATB case, so these synthetic compositions were paired with a randomly selected Bach chorale in a some- what different comparative listening task. Harmonizations were synthesized by extracting part(s) from a randomly se- lected Bach chorale and filling in the remaining parts of the composition using the method previously described in sec- tion 2.4. Original compositions (questions labelled SATB) were generated by providing a START symbol followed by ancestral sampling as previously described in section 2.3 until an END symbol is reached. The final audio provided in the questions were obtained by rendering the composi- tions using the Piano instrument from the Fluid R3 GM SoundFont.
We only considered the first response per IP address of participants who had played both choices in every question at least once and completed all five questions. This totaled 2336 participants at the time of writing, making our study one of the largest subjective listening evaluation of an au- tomatic composition system to date.
An informal analysis sug- gests that while some neurons are ambiguous to interpreta- tion, other neurons correlate significantly with recognized music-theoretic objects, particularly chords (see fig. 4). To our knowledge, this is the first reported evidence for an LSTM optimized for automatic composition learning music-theoretic concepts without explicit prior informa- tion. This invites a follow-up study testing the statistical significance of these observations.
